---
title: "Ori + Trans + Top 4 vars"
date: "`r Sys.time()`"
output: html_document
---
Get random effects coefficients and partial residuals from gamm4_2 models

```{r get utlis}
source("mmd_utils.R")
source("gamm4_utils.R")
```

```{r load_packages,message=FALSE}
# devtools::install_github("glmmTMB/glmmTMB/glmmTMB")
# devtools::install_github("bbolker/broom.mixed")
# devtools::install_github("lme4/lme4")
    # I needed to manually install 'tidyr' before to succeed with 'broom.mixed' installation
load_all_pkgs() 
theme_set(theme_bw())
zmargin <- theme(panel.spacing=grid::unit(0,"lines"))
zmarginx <- theme(panel.spacing.x=grid::unit(0,"lines"))
library('pander')  ## for tables
sapply(pkgList,function(x) format(packageVersion(x)))
comb_out <- function(p,fn,...) {
    print(p)
    htmlwidgets::saveWidget(ggplotly(p,...),fn)
}
# additional libraries
# library(plyr); library(rgdal);library(raster); library(sp); library(fields); library(plotrix)
```

```{r knitr_setup,echo=FALSE}
library(knitr)
opts_chunk$set(error=FALSE)
```


```{r source files}
load("bestmodels_gamm4.RData")
load("allfits_sum_lmer.RData") 
load("allfits_sum_gamm4.RData")
load("allfits_restr_gamm4.RData")
```


Data from Enric (includes area and lat/long coordinates):

```{r get_data}
L <- load("ecoreg.RData")
ecoreg <- ecoreg
```


#### Partial residuals:

```{r part_resid,warning=FALSE}
## fix_NAs needed for remef() results: remef_allran does it automatically
fix_NAs <- function(rem,model) {
    if (!is.null(nastuff <- attr(model.frame(model),"na.action"))) {
        return(napredict(nastuff,rem))
    } else return(rem)
}
```
  

### Plots of the original variables: 

```{r plots_ori_vars,warning=FALSE}
resplabs <- c(NPP_mean="NPP (g*C/m2*year)",
             NPP_cv_inter="CV of NPP",
             Feat_mean="Feat (% of NPP)",
             Feat_cv_inter="CV of feat")
taxvars <- c(b="mbirds",a="mamph",m="mmamm")
taxlabs <- c(mbirds="Birds",mamph="Amphibians",mmamm="Mammals")
respvars <- c("npp","nppcv","feat","featcv")
pfun <- function(taxon,respvar) {
    (ggplot(ecoreg,aes_string(x=respvar,y=taxon,colour="biome"))
            + geom_point()
            + theme(legend.position="none")
        + labs(y="N sp/(100 km2)",x=resplabs[[respvar]],title=taxlabs[[taxon]]))
}

plist <- lapply(taxvars,
       function(tax) {
    L <- lapply(names(resplabs),function(r) pfun(tax,r))
    return(setNames(L,respvars))
})
names(plist) <- names(taxvars)
grid.arrange(grobs=plist$b,ncol=2,nrow=2)
```

### Plots of the transformed variables (log and scaled)

```{r plots_trans_vars,warning=FALSE}
tbnpp <- (ggplot(ecoreg,aes(NPP_log,mbirds_log,colour=biome))
    + geom_point()
    + theme(legend.position="none") + labs(x = "log NPP", y = "log Birds")
tbnppcv <- ggplot(ecoreg,aes(NPP_cv_sc,mbirds_log,colour=biome)) + geom_point() + theme(legend.position="none") + labs(x = "scaled CV of NPP", y = "log Birds")
tbfeat <- ggplot(ecoreg,aes(Feat_log,mbirds_log,colour=biome)) + geom_point() + theme(legend.position="none") + labs(x = "log Feat", y = "log Birds")
tbfeatcv <- ggplot(ecoreg,aes(Feat_cv_sc,mbirds_log,colour=biome)) + geom_point() + theme(legend.position="none") + labs(x = "scaled CV of Feat", y = "log Birds")

tanpp <- ggplot(ecoreg,aes(NPP_log,mamph_log,colour=biome)) + geom_point() + theme(legend.position="none") + labs(x = "log NPP", y = "log Amphibians")
tanppcv <- ggplot(ecoreg,aes(NPP_cv_sc,mamph_log,colour=biome)) + geom_point() + theme(legend.position="none") + labs(x = "scaled CV of NPP", y = "log Amphibians")
tafeat <- ggplot(ecoreg,aes(Feat_log,mamph_log,colour=biome)) + geom_point() + theme(legend.position="none") + labs(x = "log Feat", y = "log Amphibians")
tafeatcv <- ggplot(ecoreg,aes(Feat_cv_sc,mamph_log,colour=biome)) + geom_point() + theme(legend.position="none") + labs(x = "scaled CV of Feat", y = "log Amphibians")

tmnpp <- ggplot(ecoreg,aes(NPP_log,mmamm_log,colour=biome)) + geom_point() + theme(legend.position="none") + labs(x = "log NPP", y = "log Mammals")
tmnppcv <- ggplot(ecoreg,aes(NPP_cv_sc,mmamm_log,colour=biome)) + geom_point() + theme(legend.position="none") + labs(x = "scaled CV of NPP", y = "log Mammals")
tmfeat <- ggplot(ecoreg,aes(Feat_log,mmamm_log,colour=biome)) + geom_point() + theme(legend.position="none") + labs(x = "log Feat", y = "log Mammals")
tmfeatcv <- ggplot(ecoreg,aes(Feat_cv_sc,mmamm_log,colour=biome)) + geom_point() + theme(legend.position="none") + labs(x = "scaled CV of Feat", y = "log Mammals")

grid.arrange(bnpp,bnppcv,bfeat,bfeatcv,tbnpp,tbnppcv,tbfeat,tbfeatcv,ncol=4,nrow=2)
```
  
  
### Plots of residuals  
  
I generated several "lists" of plots (using `plotfun`) with the lapply function following Ben's code. In each case, I changed the `xvar` and `auxvar` to plot the 4 top most variables for each taxa (aside from NPP)  

**Amph:** NPP_log:Feat_cv_sc, NPP_log:Feat_log, Feat_log:NPP_cv_sc, Feat_log  
  
**Birds:** Feat_log, NPP_log:Feat_cv_sc, NPP_cv_sc, Feat_cv_sc  
  
**Mammals:** Feat_cv_sc, Feat_log, NPP_log:Feat_cv_sc, Feat_log:NPP_cv_sc  
  

```{r }
# trying to backtrans
plotfun(best_models[[3]],respvar='mbirds_log',xvar='Feat_log',auxvar=NULL,data=ecoreg,backtrans=TRUE,log="xy")

##plotfun(best_models[[3]],respvar='mbirds_log',xvar='Feat_log',auxvar=NULL,data=ecoreg,backtrans_magic('mbirds_log','mbrids','Feat_log'))
```  
  

```{r lists_of_plots,warning=FALSE}
dd <- list(x=c("NPP_log","NPP_log","NPP_log","Feat_log",
               "NPP_cv_sc","Feat_cv_sc"),
           aux=list(NULL,"Feat_cv_sc","NPP_cv_sc",NULL,NULL,NULL))
dd$lab <- Map(function(x,v) if (is.null(v)) x else paste(x,v,sep=":"), dd$x,dd$aux)

## get partial residuals, plot
rp_fun <- function(x,v,lab) {
    lapply(best_models,
           function(m) {
        ## FIXME: for interactions, keep both + interaction;
        ##  split into facets, etc
        ecoreg$rem1 <- remef_allran(m,ecoreg,fixed_keep=lab)
        plotfun(m,xvar=x,
                respvar="rem1",auxvar=v,data=ecoreg,ylim=NULL)+
            theme(legend.position="none")+
            ggtitle(lab)
    })
}

rp <- Map(rp_fun, dd$x,dd$aux, dd$lab)
## what is wrong with intercept etc.?
pp <- plotfun(best_models[[1]],auxvar=NULL,
              ylim=c(-2.5,4))+theme(legend.position="none")
grid.arrange(rp[[1]][[1]],pp,nrow=1)

```

# **Graphical outputs** --------------------------------------------------------------------------------------
.

### Graphic example to see the color of each biome in the plots below
```{r example_plot,,warning=FALSE,fig.width=8,fig.height=6,echo=FALSE}
ggplot(ecoreg,aes(NPP_mean,mbirds,colour=biome)) + geom_point() + labs(x = "NPP (g*C/m2*year)", y = "Birds (N sp/km2)")
```
  
.
. 

## **NPP effects** for all taxa ------------------------------------------------------------------------  
  
### Original + transformed values (xy plot)    
```{r NPP_ori_trans,warning=FALSE,fig.width=10,fig.height=6.5, echo=FALSE}
grid.arrange(anpp,bnpp,mnpp,tanpp,tbnpp,tmnpp,ncol=3,nrow=2)
```


### Predicted values (plotfun)
```{r NPP_log,warning=FALSE,fig.width=10,fig.height=3.8, echo=FALSE}
grid.arrange(rp[[1]]$mamph_log,rp[[1]]$mbirds_log,
             rp[[1]]$mmamm_log,ncol=3,nrow=1)
```

  
  
## **Amphibians** ----------------------------------------------------------------------------------------
  
### Amphibians Original + transformed values (xy plot)    
```{r amph_ori,warning=FALSE,fig.width=10,fig.height=5.5, echo=FALSE}
grid.arrange(anpp,anppcv,afeat,afeatcv,tanpp,tanppcv,tafeat,tafeatcv,ncol=4,nrow=2)
```

### Amphibians 4 top most effects (plotfun)
```{r amph_plots,warning=FALSE,fig.width=10,fig.height=8, echo=FALSE}
grid.arrange(rp[[2]]$mamph_log,rp[[3]]$mamph_log,rp[[4]]$mamph_log,rp[[5]]$mamph_log,ncol=2,nrow=2)
```
  
  
## **Birds** ----------------------------------------------------------------------------------------------
  
### Birds Original + transformed values (xy plot)    
```{r birds_ori,warning=FALSE,fig.width=10,fig.height=5.5, echo=FALSE}
grid.arrange(bnpp,bnppcv,bfeat,bfeatcv,tbnpp,tbnppcv,tbfeat,tbfeatcv,ncol=4,nrow=2)
```    
    
### Birds 4 top most effects (plotfun)    
```{r birds_plots,warning=FALSE,fig.width=10,fig.height=8, echo=FALSE}
grid.arrange(rp[[5]]$mbirds_log,rp[[2]]$mbirds_log,rp[[6]]$mbirds_log,rp[[7]]$mbirds_log,ncol=2,nrow=2)
```
  
  
## **Mammals** -------------------------------------------------------------------------------------------
  
### Mammals Original + transformed values (xy plot)    
```{r mamm_ori,warning=FALSE,fig.width=10,fig.height=5.5, echo=FALSE}
grid.arrange(mnpp,mnppcv,mfeat,mfeatcv,tmnpp,tmnppcv,tmfeat,tmfeatcv,ncol=4,nrow=2)
```     
    
### Mammals 4 top mos effects (plotfun)    
```{r mamm_plots,warning=FALSE,fig.width=10,fig.height=8, echo=FALSE}
grid.arrange(rp[[7]]$mmamm_log,rp[[5]]$mmamm_log,rp[[2]]$mmamm_log,rp[[4]]$mmamm_log,ncol=2,nrow=2)
```







```{r}
m <- best_models$mbirds_log
rem1 <- remef_allran(m,ecoreg,c(fixed_keep="NPP_log","Feat_log",
                                "NPP_log:Feat_log"))
ecoreg$rem1 <- rem1
qq <- c(1/3,2/3)
## qq <- c(0.1,0.5,0.9)
ecoreg$feat_cut <- cut(ecoreg$Feat_log,
                       c(-Inf,quantile(ecoreg$Feat_log,qq),Inf))
plotfun(m,xvar="NPP_log",
        respvar="rem1",
        auxvar=NULL,
        ##auxvar="Feat_log",
        data=ecoreg,ylim=c(-1,0.5))+
    theme(legend.position="none")+
    facet_wrap(~feat_cut,labeller=label_both)
## geom_point(aes(size=Feat_log,y=rem1,colour=biome,shape=flor_realms))+
##    scale_size_continuous(range=c(0.1,8))
    


