---
title: "diversity analyses with brm/full data set etc."
---

```{r pkgs}
source("mmd_utils.R")
source("gamm4_utils.R")
L <- load("ecoreg.RData")
library(broom)
library(broom.mixed)
library(stringr)
library(dplyr)
library(brms)
library(corrplot)
library(dplyr)
library(ggplot2); theme_set(theme_bw())
```


## getting brms to run (original data set)

`brms` (actually the underlying `Stan` platform may complain about "divergent steps". There is quite a bit of discussion about how to fix this, e.g. [brms github issues](https://github.com/paul-buerkner/brms/issues/128), but it seems that using `adapt_delta=0.99` takes care of problems that we have.

from `?set_prior`:

- default prior on SDs is half-Student with 3 df, scale parameter based on std dev of response (>=10) e.g. `set_prior("<prior>", class="sd", group="<group>")` (is this going to regularize enough??)
- default correlation priors `lkj_(eta)`: `eta=1` - flat; `eta > 1` extreme correlations less likely

Looks like the default priors might be OK.

## basic results

```{r}
L <- load("allfits_sum_gamm4.RData") ## 4 taxa x 27 fits, using gamm4
L <- load("brms_batch.RData")
gamm4_coefs <- (
  gamm4_res$coef
    %>% inner_join(filter(gamm4_res$sum,best))
    %>% select(-c(AIC,singular,df,best)))
## 11 "non-varying"
## 375 "varying"
## 45 "hierarchical"
tt0 <- rbind(tidy(b1,par_type="hierarchical"),
             tidy(b1,par_type="non-varying"))
unique(tt0$term)
unique(gamm4_coefs$term)
tt1 <- (tt0
    %>% mutate(term=str_replace(term,"Intercept","(Intercept)"),
               ## https://stackoverflow.com/questions/35804379/r-stringr-str-extract-how-to-do-postive-lookbehind/35804399
               ## extract group from term name. FRAGILE if group names contain "_" ...
               grp=str_match(term,"^sd_([^_]+)")[,2])
                         
    %>% mutate(cat=case_when(grepl("^cor",term) ~ "cor",
                             grepl("^sd",term) ~ "sd",
                             TRUE ~ "fixed"),
               ## reorder
               cat=factor(cat,levels=c("fixed","sd","cor")))
    %>% filter(term!="Intercept")
    ## reorder term by estimate *within* categories -- fussy ...
    %>% arrange(cat,estimate)
    %>% mutate(term=factor(term,levels=term))
)
ggplot(tt1,aes(estimate,term))+
    ggstance::geom_pointrangeh(aes(xmin=lower,xmax=upper))+
    facet_wrap(~cat,scale="free",ncol=1)+
    geom_vline(xintercept=0,lty=2)
```

What do we conclude?

```{r}
vv <- VarCorr(b1)
## vv$biome$cor[,"Estimate",]
## vv$biome$cor[,"Est.Error",]
```

## Markov random field stuff

- `corsar(W, type=c("lag","error"))`; "lag" = SAR on response values; "error" = SAR on residuals
- [CrossValidated](https://stats.stackexchange.com/questions/277/spatial-statistics-models-car-vs-sar)
     - non-spatial: "My House Value is a function of my home Gardening Investment"
     - SAR: "My House Value is a function of the House Values of my neighbours."
     - CAR: "My House Value is a function of the Gardening Investment of my neighbours"
	 - [issue describing mrf smooth](https://github.com/paul-buerkner/brms/issues/128)
	 - `?mgcv::mrf`: can give polygons (`polys`), or neighborhood structure (`nb`)

