---
title: "diversity analyses with brm/full data set etc."
---

```{r pkgs,message=FALSE}
source("mmd_utils.R")
source("gamm4_utils.R")
L <- load("ecoreg.RData")
library(broom.mixed)
library(stringr)
library(dplyr)
library(brms)
library(corrplot)
library(dplyr)
library(ggplot2); theme_set(theme_bw())
```

```{r opts}
knitr::opts_chunk$set(echo=FALSE)
```

## getting brms to run (original data set)

`brms` (actually the underlying `Stan` platform may complain about "divergent steps". There is quite a bit of discussion about how to fix this, e.g. [brms github issues](https://github.com/paul-buerkner/brms/issues/128), but it seems that using `adapt_delta=0.99` takes care of problems that we have.

from `?set_prior`:

- default prior on SDs is half-Student with 3 df, scale parameter based on std dev of response (>=10) e.g. `set_prior("<prior>", class="sd", group="<group>")` (is this going to regularize enough??)
- default correlation priors `lkj_(eta)`: `eta=1` - flat; `eta > 1` extreme correlations less likely

Looks like the default priors might be OK.

## basic results

```{r arrange,warning=FALSE}
## FIXME: eliminate warning in bind_rows by making factors <-> characters
L <- load("allfits_sum_gamm4.RData") ## 4 taxa x 27 fits, using gamm4
L <- load("allfits_sum_brms.RData")
gamm4_coefs <- (
    gamm4_res$coef
    ## take just best model
    %>% inner_join(filter(gamm4_res$sum,best),by=c("taxon","model"))
    %>% select(-c(AIC,singular,df,best)))

## remove fixed-effect intercept (only)
noint <- . %>% filter(!grepl("^.?Intercept.?$",term))
## unique(brms_res$coefs$term)
## unique(gamm4_coefs$term)
brms_res$coefs <- (brms_res$coefs
    %>% noint()
    ## replace missing parens
    %>% mutate(term=str_replace(term,"Intercept","(Intercept)"),
               term=str_replace(term,"logarea_km2","log(area_km2)"))
    ## don't want intercept anyway ...
    %>% filter(!grepl("^.?Intercept.?$",term))
    %>% mutate(cat=case_when(grepl("^cor",term) ~ "cor",
                             grepl("^sd",term) ~ "sd",
                             TRUE ~ "fixed"),
               ## reorder
               cat=factor(cat,levels=c("fixed","sd","cor")))
    ## reorder term by estimate *within* categories -- fussy ...
    %>% arrange(cat,estimate)
    %>% mutate(term=factor(term,levels=unique(term)))
)
```

Random-effects coefficients:

```{r brms_all_coefs,fig.width=10,fig.height=8}
## FIXME: adjust vertical spacing?
## https://stackoverflow.com/questions/31572239/set-space-in-facet-wrap-like-in-facet-grid
cc <- (brms_res$coefs
    %>% filter(cat != "fixed", group!="Residual"))
g0 <- ggplot(cc,aes(estimate,term,colour=taxon,shape=taxon))+
    ggstance::geom_pointrangeh(aes(xmin=conf.low,xmax=conf.high))+
    facet_grid(.~group,scale="free")+
    geom_vline(xintercept=0,lty=2)+
    labs(y="")+
    scale_colour_brewer(palette="Set1")
res <- list()
for (i in c("sd","cor")) {
    res[[i]] <- g0 %+% filter(cc,cat==i)
}
cowplot::plot_grid(plotlist=res,ncol=1,align="v")
```

Comparing `brms` models to "best" `gamm4` model:

```{r arrange2, warning=FALSE}
comb <- bind_rows(brms=brms_res$coefs,
                  gamm4=noint(gamm4_coefs),.id="platform")
comb_fixed <- (comb
    %>% filter(effect=="fixed")
    %>% mutate(term=reorder(term,estimate))
)
```

```{r brms_gamm4_comp,fig.width=9,fig.height=7}
pd <- ggstance::position_dodgev(height=0.35)
ggcom <- ggplot(comb_fixed,aes(estimate,term,colour=platform,shape=platform))+
    scale_colour_brewer(palette="Dark2")+
    ggstance::geom_pointrangeh(aes(xmin=conf.low,xmax=conf.high),
                               size=0.5,
                               position=pd)+
    facet_wrap(~taxon,scale="free")+
    geom_vline(xintercept=0,lty=2)
print(ggcom)
```

Birds only
```{r}
ggcom %+% filter(comb_fixed,taxon=="mbirds_log")
```

What do we conclude? A few confidence intervals are wider (especially, `NPP_log` for `plants_log`)

## plotting functions

Do they work with `brms`? How much adjustment is required?

```{r}
L <- load("allfits_brms.RData")
plotfun(allfits_brms[[1]])
plotfun(allfits_brms[[1]],backtrans=TRUE,log="xy")
```

## Spatial


## Markov random field stuff

- `corsar(W, type=c("lag","error"))`; "lag" = SAR on response values; "error" = SAR on residuals
- [CrossValidated](https://stats.stackexchange.com/questions/277/spatial-statistics-models-car-vs-sar)
     - non-spatial: "My House Value is a function of my home Gardening Investment"
     - SAR: "My House Value is a function of the House Values of my neighbours."
     - CAR: "My House Value is a function of the Gardening Investment of my neighbours"
	 - [issue describing mrf smooth](https://github.com/paul-buerkner/brms/issues/128)
	 - `?mgcv::mrf`: can give polygons (`polys`), or neighborhood structure (`nb`)

