---
title: exploring `gamm4` residuals
date: "`r Sys.time()`"
---


The fitted model objects returned by `gamm4` are lists of two components, called `mer` and `gam`. How do we get the predictions/residuals we want, i.e. incorporating effects of all three of (1) regular fixed effect terms, (2) smooth terms, (3) random effect terms?

**tl;dr** `residuals(m1$mer)` should be fine.

```{r pkgs, message=FALSE}
##
library(gamm4)
library(lme4)     ## for simulate()
library(splines)  ## for ns()
library(ggplot2); theme_set(theme_bw())
library(cowplot)  ## for plot_grid()
```

Set up simulated data with

- regular fixed effect (predictor is 0/1, so this should be pretty easy to see)
- smooth term (10-point spline)
- random effect (scalar/intercept-only model)

```{r sim}
set.seed(101)
dd <- data.frame(z=rep(0:1,each=250),
                 x=rnorm(500),f=factor(rep(1:10,50)))
## check number of parameters, order:
colnames(model.matrix(~z+ns(x,10),data=dd))
## spline parameters (pretty arbitrary - we need 10 values)
## won't correspond exactly to parameters from mgcv::gam anyway,
##  as it uses a different spline framework
sparms <- rep(c(1,0),5)
## simulate data:
## (suppress message about unnamed theta/beta)
dd$y <- suppressMessages(simulate(~z+ns(x,10)+(1|f),
                  newdata=dd,
                  newparams=list(beta=c(1,1,sparms),
                                 ## random-effects sd (scaled by sigma)
                                 theta=30,
                                 ## small residual error so we can see
                                 ##  what's going on clearly
                                 sigma=0.01),
                  family=gaussian))[[1]]
## order by x for plotting convenience
dd <- dd[order(dd$x),]
```

See what we've got:

```{r gg0}
gg0 <- ggplot(dd,aes(x,y,colour=f))+geom_point()+
    geom_line(aes(group=interaction(z,f)))+
    scale_colour_discrete(guide=FALSE)
print(gg0)
```

Fit the `gamm4` model:
```{r fit}
## n.b. need parentheses around random effect
m1 <- gamm4(y~z+s(x), random=~(1|f),data=dd)
```

Check parameters:
```{r parms}
fixef(m1$mer) ## includes z + a mystery parameter (smooth effect?)
coef(m1$gam)  ## includes z + smooth terms
```

Look at predicted from each component vs observed:
neither prediction captures everything?

```{r pred_obs,fig.width=8}
par(mfrow=c(1,2))
plot(dd$y,predict(m1$mer),xlab="obs",ylab="pred",main="mer pred")
abline(a=0,b=1,col="red")
plot(dd$y,predict(m1$gam),xlab="obs",ylab="pred",main="gam pred")
abline(a=0,b=1,col="red")
```

But adding random effects to gam prediction by hand gives
mer prediction ...

```{r pred_aug}
re <- ranef(m1$mer)$f[,1]  ## random-effects vector
augpred <- c(predict(m1$gam)+ re[dd$f])
all.equal(augpred,predict(m1$mer))
```

... suggesting that mer prediction is what we need

This model (which I originally fitted) *would* capture everything almost exactly
but slow (lots of FE parameters):
```{r ns_fit,eval=FALSE}
gamm4(y~z+ns(x,10), random=~(1|f),data=dd)
```

If we plot the predicted rather than the actual data, we clearly
see that all of the components of the pattern (fixed effect,
smooth terms, random terms) are incorporated in the prediction:

```{r plot_res}
plot_grid(gg0+ggtitle("sim data"),
          gg0 %+% transform(dd,y=predict(m1$mer)) + ggtitle("mer predictions"))
```


So `residuals(m1$mer)` should be fine:

```{r resids}
all.equal(residuals(m1$mer),dd$y-predict(m1$mer))
```

See also: https://stat.ethz.ch/pipermail/r-sig-mixed-models/2011q2/013602.html
