---
title: "Original, transformed, & partial residuals plots"
date: "`r Sys.time()`"
output:
  html_document:
    code_folding: hide
---

**FIXME:** 

- figure out amphibian offset ...
     - what is different about amphibians? Fixed potential NA issues, I think
	 - diversities are lower than for birds/mammals (also checked plants, they have even higher diversities and look fine)
	 - don't see how it could be a nonlinear-averaging thing (we do all computations on the log scale and display the results on the log scale) but low-diversity could be where such effects are strongest
	 - could it be the effects of the smooth? (i.e. value at average x,y rather than leaving out the smooth term entirely) (TEST: check whether this effect occurs for lme4 models as well)
- two-line axis labels?
- facet/strip redundant axes and labels?
- prettier titles?
  
```{r get_utils}
source("utils.R")
source("gamm4_utils.R")
```

```{r load_packages,message=FALSE}
load_all_pkgs()
graphics_setup()
library('pander')  ## for tables
sapply(pkgList,function(x) format(packageVersion(x)))
```

```{r knitr_setup,echo=FALSE}
library(knitr)
opts_chunk$set(error=FALSE)
```


```{r source files}
best_models <- readRDS("bestmodels_gamm4.rds")
best_lmer_models <- readRDS("allfits_lme4.rds")
ecoreg <- readRDS("ecoreg.rds")
```

### Plots of the original variables: 

```{r plots_ori_vars,warning=FALSE}
taxon_names <- c("mbirds","mamph","mmamm","plants")
pretty_taxon_names <- c("Birds","Amphibians","Mammals","Plants")
respvars <- paste0(taxon_names,"_log")
predvars <- c("NPP_log_sc","Feat_log_sc","NPP_cv_sc","Feat_cv_sc")
## plot raw values of diversity for a given taxon/predictor
plotfun(model=NULL, respvar="mbirds_log",
        xvar="NPP_log_sc",
        backtrans=TRUE,log="xy")
## generate all combinations
orig_plots <- list()
for (i in seq_along(taxon_names)) {
    orig_plots[[taxon_names[i]]] <- list()
    for (pp in predvars) {
      orig_plots[[taxon_names[i]]][[pp]] <-
        plotfun(model=NULL,respvar=respvars[i],
                xvar=pp,data=ecoreg,
                backtrans=TRUE,log="xy")+
                theme(legend.position="none")
    }
}
plot_grid(plotlist=orig_plots[["mbirds"]],align="hv",nrow=2)
```

### Plots of the transformed variables (log and scaled)

```{r plots_trans_vars,warning=FALSE}
## names are the names of the variable in the data set;
##  the character strings are the names to put on the axis
## this time using log-scaled etc.
pred_names <- c(NPP_mean="NPP g*C/m2*year",
                Feat_mean="Feat (% of NPP)",
                NPP_cv_inter="CV of NPP",
                Feat_cv_inter = "CV of Feat")
sc_taxon_names <- paste(pretty_taxon_names, "N sp/km2")
names(sc_taxon_names) <- paste0(taxon_names,"_log")
sc_pred_names <- c(NPP_log_sc="log NPP g*C/m2*year",
                   Feat_log_sc="log Feat (% of NPP)",
                   NPP_cv_sc="scaled CV of NPP",
                   Feat_cv_sc = "scaled CV of Feat")
do_plots_1 <- function(taxon="mbirds",pred="NPP_mean",
                       s_names=taxon_names,
                       p_names=pred_names) {
    ggplot(ecoreg,aes_string(x=pred,y=taxon,colour="biome")) +
        geom_point() +
        theme(legend.position="none") +
        labs(y=s_names[taxon], x=p_names[pred])
}
## do all combinations
sc_plots <- list()
for (tt in names(sc_taxon_names)) {
    sc_plots[[tt]] <- list()
    for (pp in names(sc_pred_names)) {
        sc_plots[[tt]][[pp]] <- do_plots_1(tt,pp,
                                           p_names=sc_pred_names,
                                           s_names=sc_taxon_names)
    }
}
## for a given taxon, draw unscaled plots in row 1 and scaled plots in row 2
both_plots <- function(taxon) {
    plot_grid(plotlist=c(orig_plots[[taxon]],
                         sc_plots[[paste0(taxon,"_log")]]),
              nrow=2,
              align="hv")
}
```

### Plots of residuals  
  
I generated several "lists" of plots (using `plotfun`) with the lapply function following Ben's code. In each case, I changed the `xvar` and `auxvar` to plot the 4 top most variables for each taxa (aside from NPP)  

- **Amph:** `NPP_log:Feat_cv_sc`, `NPP_log:Feat_log`, `Feat_log:NPP_cv_sc`, `Feat_log`
- **Birds:** `Feat_log`, `NPP_log:Feat_cv_sc`, `NPP_cv_sc`, `Feat_cv_sc`
- **Mammals:** `Feat_cv_sc`, `Feat_log`, `NPP_log:Feat_cv_sc`, `Feat_log:NPP_cv_sc`
  
```{r backtrans}
taxon <- "mbirds_log"
plotfun(best_models[[taxon]],
        respvar=taxon,
        xvar='Feat_log_sc',auxvar=NULL,data=ecoreg,backtrans=TRUE,log="xy")
```
  
```{r lists_of_plots,warning=FALSE}
## partial residuals plots
remef_plot <- function(taxon="mbirds_log",xvar="NPP_log",
                       auxvar=NULL,title=NULL) {
    m <- best_models[[taxon]]
    if (is.null(title)) {
        title <- if (is.null(auxvar)) xvar else {
             paste(xvar,auxvar,sep=":")                                     
                                              }
    }
    pp <- (plotfun(m,xvar=xvar,respvar="partial_res",
                   auxvar=auxvar,data=ecoreg,
                   backtrans=TRUE,log="xy") 
      + theme(legend.position="none")
      + ggtitle(title)
    )
    return(pp)
}
## rp1: NPP_log
## rp2: NPP_log:Feat_cv_sc
## rp3: NPP_log:Feat_log
## rp4: Feat_log:NPP_cv_sc
## rp5: Feat_log
## rp6: NPP_cv_sc
## rp7: Feat_cv_sc
rem_predvars <- list("NPP_log_sc",
                     "NPP_log_sc",
                     "NPP_log_sc",
                     "Feat_log_sc",
                     "Feat_log_sc",
                     "NPP_cv_sc",
                     "Feat_cv_sc")
rem_auxvars <- list(NULL,"Feat_cv_sc","Feat_log_sc",
                    "NPP_cv_sc",NULL,NULL,NULL)
## construct all combinations of partial residuals for each taxon
remef_plots <- list()
for (tt in names(sc_taxon_names)) { ## for each taxon
    remef_plots[[tt]] <- list()
    for (pp in seq_along(rem_predvars)) { ## for each predictor variable
      ## cat(tt,pp,"\n")
      ## cat(tt," ",pp,"\n")
      ## generate and save the partial residuals plot
      remef_plots[[tt]][[pp]] <- remef_plot(tt,xvar=rem_predvars[[pp]],
                                            auxvar=rem_auxvars[[pp]])
      ## print(remef_plots[[tt]][[pp]])
    }
}
```

# Graphical outputs


### Graphic example to see the color of each biome in the plots below

```{r example_plot,,warning=FALSE,fig.width=8,fig.height=6}
ggplot(ecoreg,aes(NPP_mean,mbirds,colour=biome)) + geom_point() + labs(x = "NPP (g*C/m2*year)", y = "Birds (N sp/km2)")
```
  
## Amphibians
  
### Original + transformed values (xy plot)    
```{r amph_ori,warning=FALSE,fig.width=10,fig.height=5.5}
both_plots("mamph")
```

### Top 4 effects
```{r amph_plots,warning=FALSE,fig.width=10,fig.height=8}
do.call(grid.arrange,
        c(remef_plots[["mamph_log"]][c(2,3,4,5)],
          ## NPP_log:Feat_csv, NPP_log:Feat_log, Feat_log:NPP_cv_sv, Feat_log
          list(ncol=2)))
```
    
## Birds
  
### Original + transformed values
```{r birds_ori,warning=FALSE,fig.width=10,fig.height=5.5}
both_plots("mbirds")
```    
    
### Top 4 effects
	
```{r birds_plots,warning=FALSE,fig.width=10,fig.height=8}
do.call(grid.arrange,
        c(remef_plots[["mbirds_log"]][c(5,2,6,7)],
               list(ncol=2)))
```

## Mammals

### Original + transformed values

```{r mamm_ori,warning=FALSE,fig.width=10,fig.height=5.5}
both_plots("mmamm")
```     
    
### Top 4 effects

```{r mamm_plots,warning=FALSE,fig.width=10,fig.height=8}
do.call(grid.arrange,
        c(remef_plots[["mmamm_log"]][c(7,5,2,4)],
          list(ncol=2)))
```

## questions

* why do fitted model lines look bad for amphibians? NA values/different mean? Amphibians are indeed the only data set with any NA values ...

```{r calc_na}
nna <- colSums(is.na(ecoreg))
nna[nna>0]
```

```{r calc_means}
ee <- subset(ecoreg, select=c(Area,mamph_log,mbirds_log,
                              NPP_mean,NPP_cv_inter,Feat_mean,Feat_cv_inter))
predvars <- c("NPP_mean","NPP_cv_inter","Feat_mean","Feat_cv_inter","Area")
rbind(amph=colMeans(ee[!is.na(ee$mamph_log),predvars]),
      birds=colMeans(ee[!is.na(ee$mbirds_log),predvars]))
```

Means are indeed *slightly* different, but is this enough to drive the observed difference?

