---
title: "Original, transformed, & partial residuals plots"
date: "`r Sys.time()`"
output:
  html_document:
    code_folding: hide
---

```{r get utlis}
source("mmd_utils.R")
source("gamm4_utils.R")
```

```{r load_packages,message=FALSE}
load_all_pkgs() 
theme_set(theme_bw())
zmargin <- theme(panel.spacing=grid::unit(0,"lines"))
zmarginx <- theme(panel.spacing.x=grid::unit(0,"lines"))
library('pander')  ## for tables
sapply(pkgList,function(x) format(packageVersion(x)))
comb_out <- function(p,fn,...) {
    print(p)
    htmlwidgets::saveWidget(ggplotly(p,...),fn)
}
```

```{r knitr_setup,echo=FALSE}
library(knitr)
opts_chunk$set(error=FALSE)
```


```{r source files}
load("bestmodels_gamm4.RData")
load("allfits_sum_lmer.RData") 
load("allfits_sum_gamm4.RData")
load("allfits_restr_gamm4.RData")
```


Data from Enric (includes area and lat/long coordinates):

```{r get_data}
L <- load("ecoreg.RData")
ecoreg <- ecoreg
```

### Plots of the original variables: 

```{r plots_ori_vars,warning=FALSE}
pred_names <- c(NPP_mean="NPP g*C/m2*year",
                NPP_cv_inter="CV of NPP",
                Feat_mean="Feat (% of NPP)",
                Feat_cv_inter = "CV of Feat")
taxon_names <- paste(c("Birds","Amphibians","Mammals"),
                    "N sp/km2")
names(taxon_names) <- c("mbirds","mamph","mmamm")
do_plots_1 <- function(taxon="mbirds",pred="NPP_mean",
                       s_names=taxon_names,
                       p_names=pred_names) {
    ggplot(ecoreg,aes_string(x=pred,y=taxon,colour="biome")) +
        geom_point() +
        theme(legend.position="none") +
        labs(y=s_names[taxon], x=p_names[pred])
}
orig_plots <- list()
for (tt in names(taxon_names)) {
    orig_plots[[tt]] <- list()
    for (pp in names(pred_names)) {
        orig_plots[[tt]][[pp]] <- do_plots_1(tt,pp)
    }
}
do.call(grid.arrange,c(orig_plots[["mbirds"]],list(nrow=2)))
```

### Plots of the transformed variables (log and scaled)

```{r plots_trans_vars,warning=FALSE}
sc_pred_names <- c(NPP_log="log NPP g*C/m2*year",
                   NPP_cv_sc="scaled CV of NPP",
                   Feat_log="log Feat (% of NPP)",
                   Feat_cv_sc = "scaled CV of Feat")

sc_taxon_names <- paste("log",taxon_names)
names(sc_taxon_names) <- paste0(names(taxon_names),"_log")

sc_plots <- list()
for (tt in names(sc_taxon_names)) {
    sc_plots[[tt]] <- list()
    for (pp in names(sc_pred_names)) {
        sc_plots[[tt]][[pp]] <- do_plots_1(tt,pp,
                                           p_names=sc_pred_names,
                                           s_names=sc_taxon_names)
    }
}
both_plots <- function(taxon) {
    do.call(grid.arrange,c(orig_plots[[taxon]],
                           sc_plots[[paste0(taxon,"_log")]],
                           list(nrow=2)))
}
```

FIXME: y-axis names
  
### Plots of residuals  
  
I generated several "lists" of plots (using `plotfun`) with the lapply function following Ben's code. In each case, I changed the `xvar` and `auxvar` to plot the 4 top most variables for each taxa (aside from NPP)  

**Amph:** NPP_log:Feat_cv_sc, NPP_log:Feat_log, Feat_log:NPP_cv_sc, Feat_log  
  
**Birds:** Feat_log, NPP_log:Feat_cv_sc, NPP_cv_sc, Feat_cv_sc  
  
**Mammals:** Feat_cv_sc, Feat_log, NPP_log:Feat_cv_sc, Feat_log:NPP_cv_sc  
  
```{r backtrans}
plotfun(best_models[[3]],respvar='mbirds_log',xvar='Feat_log',auxvar=NULL,data=ecoreg,backtrans=TRUE,log="xy")
```  
  

```{r lists_of_plots,warning=FALSE}
remef_plot <- function(taxon="mbirds_log",predvar="NPP_log",
                       auxvar=NULL,title=NULL) {
    m <- best_models[[taxon]]
    if (is.null(title)) {
        title <- if (is.null(auxvar)) predvar else {
             paste(predvar,auxvar,sep=":")                                     
                                              }
    }
    rem1 <- remef_allran(m,ecoreg,
                         set_other="zero",
                         keep_intercept=TRUE,
                         fixed_keep=predvar)
    if (length(rem1)==nrow(ecoreg)) {
        ## if it worked ...
        ecoreg$rem1 <- rem1
        pp <- plotfun(m,xvar=predvar,respvar="rem1",
                      auxvar=auxvar,data=ecoreg,ylim=NULL)+
            theme(legend.position="none")+
            ggtitle(title)
    }
    return(pp)
}

## rp1: NPP_log
## rp2: NPP_log:Feat_cv_sc
## rp3: NPP_log:Feat_log
## rp4: Feat_log:NPP_cv_sc
## rp5: Feat_log
## rp6: NPP_cv_sc
## rp7: Feat_cv_sc
rem_predvars <- list("NPP_log",
                     "NPP_log",
                     "NPP_log",
                     "Feat_log",
                     "Feat_log",
                     "NPP_cv_sc",
                     "Feat_cv_sc")
rem_auxvars <- list(NULL,"Feat_cv_sc","Feat_log",
                    "NPP_cv_sc",NULL,NULL,NULL)

remef_plots <- list()
for (tt in names(sc_taxon_names)) {
    remef_plots[[tt]] <- list()
    for (pp in seq_along(rem_predvars)) {
        ## cat(tt," ",pp,"\n")
        remef_plots[[tt]][[pp]] <- remef_plot(tt,predvar=rem_predvars[[pp]],
                                              auxvar=rem_auxvars[[pp]])
    }
}
```

.

# **Graphical outputs** -
.
### Graphic example to see the color of each biome in the plots below
```{r example_plot,,warning=FALSE,fig.width=8,fig.height=6,echo=FALSE}
ggplot(ecoreg,aes(NPP_mean,mbirds,colour=biome)) + geom_point() + labs(x = "NPP (g*C/m2*year)", y = "Birds (N sp/km2)")
```
  
## **Amphibians**
  
### Original + transformed values (xy plot)    
```{r amph_ori,warning=FALSE,fig.width=10,fig.height=5.5, echo=FALSE}
both_plots("mamph")
```

### Top 4 effects
```{r amph_plots,warning=FALSE,fig.width=10,fig.height=8, echo=FALSE}
do.call(grid.arrange,
        c(remef_plots[["mamph_log"]][c(2,3,4,5)],
               list(ncol=2)))
```
    
## **Birds** 
  
### Original + transformed values
```{r birds_ori,warning=FALSE,fig.width=10,fig.height=5.5, echo=FALSE}
both_plots("mbirds")
```    
    
### Top 4 effects
	
```{r birds_plots,warning=FALSE,fig.width=10,fig.height=8, echo=FALSE}
do.call(grid.arrange,
        c(remef_plots[["mbirds_log"]][c(5,2,6,7)],
               list(ncol=2)))
```
## **Mammals** 
### Original + transformed values
```{r mamm_ori,warning=FALSE,fig.width=10,fig.height=5.5, echo=FALSE}
both_plots("mmamm")
```     
    
### Top 4 effects
```{r mamm_plots,warning=FALSE,fig.width=10,fig.height=8, echo=FALSE}
do.call(grid.arrange,
        c(remef_plots[["mmamm_log"]][c(7,5,2,4)],
               list(ncol=2)))
```









